% adapted by WS from SSR's Word document and from the 5-part series at
% https://www.overleaf.com/learn/latex/How_to_Write_a_Thesis_in_LaTeX_(Part_1):_Basic_Structure
% reviewed by MAM

\documentclass[12pt,twosided]{report}

\usepackage[titletoc]{appendix} % for adding appendix to TOC
\usepackage{biblatex}  % reference management
\usepackage{geometry}  % better margins and margin control
\usepackage{graphicx}  % to include figures
\usepackage{hyperref}  % internal and external links
\usepackage[utf8]{inputenc}  % support for non-ASCII characters
\usepackage{listings}  % typset code
\usepackage{outlines}  % easy nesting of lists
\usepackage{tabularx}  % more control over table column width
\usepackage{titlesec}  % customize chapter title 
\usepackage{upquote}  % prevent mishandling of single quotes in listings
\usepackage{float}
% \addbibresource{references.bib}


% TODO: Customize the appearance of hyperref links using \hypersetup
% See https://en.wikibooks.org/wiki/LaTeX/Hyperlinks#Customization

% Custom format of chapter title.
\titleformat{\chapter}[hang]{\bf\huge}{\thechapter.}{2pc}{}

% Separate folder for images named 'images'.
\graphicspath{ {images/} }

% Separate file for references
\addbibresource{references.bib}

% Replace with your title
\title{Rahguzar}

% Allow recalling document title
% from https://tex.stackexchange.com/a/15806/44301
\makeatletter\let\Title\@title\makeatother  

\begin{document}

\include{titlepage}  % title page.
\include{approval}  % approval page.

\chapter*{Dedication}
To our parents, whose unwavering love, endless sacrifices, and heartfelt prayers have been the foundation of everything we’ve accomplished. Your belief in us, even when we doubted ourselves, gave us the strength to move forward. This journey, and all that it has brought us, is a reflection of your dedication and support.

To our families and loved ones, thank you for your patience, encouragement, and constant presence. Your understanding made room for our long hours, tight deadlines, and moments of exhaustion.

To our teachers and mentors, we are deeply grateful for your guidance and the knowledge you’ve so generously shared. Your trust in our capabilities challenged us to think bigger and work harder.

To our teammates, thank you for the long nights, the shared goals, and the spirit of collaboration that brought this project to life. Rahguzar is a product of shared vision, hard work, and mutual respect.

Above all, we dedicate this work to Allah (SWT), the source of all knowledge and strength. Without His mercy and guidance, none of this would have been possible.



\chapter*{Acknowledgements}
Alhamdulillah, all praise is due to Allah (SWT), whose endless mercy, guidance, and blessings gave us the strength and clarity to complete this project.

We are sincerely grateful to Dr. Syeda Saleha Raza, from the Faculty of Computer Science at Habib University, for her academic mentorship, thoughtful guidance, and unwavering support throughout the course of this project. Her insights and encouragement consistently pushed us to approach our work with both depth and discipline.

We also extend our heartfelt thanks to Ms. Fatima Alvi, Management Trainee at SalesFlo Pvt Ltd, for her practical guidance, domain expertise, and continuous availability. Her support helped us navigate the real-world complexities of the problem we set out to solve.

We are thankful to the Computer Science faculty at Habib University for equipping us with the knowledge and skills that formed the backbone of this project. Every course, discussion, and challenge over the past four years played a role in preparing us for this journey.

We deeply appreciate the collaboration with SalesFlo Pvt Ltd, our industry partner, for providing access to relevant datasets, business context, and timely feedback. Your involvement helped us ground our work in real operational needs.

Our gratitude also goes to the Kaavish Committee and the Office of Undergraduate Research for enabling a structured, well-supported project experience and for providing necessary resources and checkpoints along the way.

To our families and friends, thank you for your constant encouragement, patience, and understanding through all the late nights, deadlines, and moments of stress. Your support made all the difference.

Finally, we acknowledge one another. As teammates, we shared a vision, faced challenges, and built something we are proud of. Rahguzar is a result of that collective effort and commitment.



\chapter*{Abstract}
Abstract goes here

% The following are automatically populated by LaTeX \chapter, \section and related, \figure, and \table.
\tableofcontents
\listoffigures
\listoftables

\chapter{Introduction}
\label{chap:intro}
\input{chapters/introduction}

\chapter{Literature Review}
\label{chap:lit}
\input{chapters/review}

\chapter{Software Requirement Specification (SRS)}
\label{chap:srs}
\input{chapters/srs}

\chapter{Software Design Specification (SDS)}
\label{chap:sds}
\input{chapters/sds}

\chapter{Methodology}
\label{chap:method}
\input{chapters/methodology}

\chapter{Development}
\label{chap:dev}
\input{chapters/development}

\chapter{Experiments and Results}
\label{chap:results}
\input{chapters/results}

\chapter{Conclusion and Future Work}
\label{chap:outro}
\input{chapters/conclusion}

\chapter{Reflections}
\label{chap:reflections}
\input{chapters/reflections}

\begin{appendices}

\titleformat{\chapter}[hang]{\bf\huge}{Appendix \thechapter.}{2pc}{}
  
% This appendix is optional.
\chapter{More Math}
\input{chapters/appendix-math}

% This appendix is required if the data set is not fully described in the main text.
\chapter{Data}
\input{chapters/appendix-data}

% This appendix is required if the code is not fully described in the main text.
\chapter{Code}
\input{chapters/appendix-code}
\end{appendices}

% Print the bibliography with a ToC entry and titled, "References".
\printbibliography[heading=bibintoc,title={References}]

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

