The Rahguzar project was conceptualized and developed in response to the growing operational challenges faced in manual PJP planning. Through a modular, heuristic-driven methodology combining clustering, scheduling, and routing optimizations, Rahguzar demonstrated substantial improvements in route efficiency, workload distribution, and scalability. The following sections summarize the project’s key outcomes, identify its limitations, and propose directions for future research and system enhancements to further advance intelligent journey planning.

\section{Conclusion}
In this project, we presented Rahguzar, a comprehensive solution for optimizing Permanent Journey Plan (PJP) scheduling in the Fast-Moving Consumer Goods (FMCG) industry. Our system integrates customer clustering, schedule generation via Evolutionary Algorithms, and route optimization using Ant Colony Optimization (ACO), delivering a robust end-to-end pipeline for sales team planning and management.

Through the integration of real-world datasets and a user-friendly web-based interface, Rahguzar enables dynamic and efficient route planning, helping sales managers minimize travel costs while maximizing coverage and visit compliance. Performance evaluations demonstrate the system’s ability to produce realistic, optimized schedules with significantly improved route efficiency compared to manually designed plans.

The modular architecture allows easy adaptation to various geographical scales, constraints, and business-specific requirements. Moreover, our backend pipeline has been successfully deployed and integrated with a React-based frontend, creating a seamless experience for end users.
\section{Future Work}
In future iterations of Rahguzar, several enhancements can be explored to improve its effectiveness and integration with real-world FMCG workflows. One key direction is the incorporation of real-time traffic data and dynamic scheduling capabilities, enabling the system to adapt routes on the fly in response to disruptions such as delays or cancellations. Integration with Salesflo’s existing mobile platform—which is already widely used by field agents—can also be explored to enable seamless access to route plans, check-ins, and live updates. This would allow on-ground activities to be synchronized with the backend optimization engine, enhancing both usability and data flow. Another promising direction is the use of historical and real-time sales data to drive scheduling decisions. By aligning visit frequency and prioritization with customer-specific sales performance, seasonal trends, and product movement, Rahguzar can help optimize not only travel efficiency but also commercial outcomes. Together, these enhancements can transform Rahguzar into a fully adaptive, data-driven system for sales force automation.